---
title: "Vietnam Leeches using Alex Crampton-Platt's 20180618 Protax run"
author: "Douglas Yu"
date: "`r Sys.Date()`"
output: html_document
---

This protax run was run using the pipeline by Axtner et al., which starts with the Midori database and set high priors for Vietnam species.  The OTU table and OTU representative seqs are in the folder:  Vietnam_leeches_swarm_lulu_20180404, and the Protax outputs are in folder Vietnam_leeches_swarm_lulu_20180404/doug_vietnam_16S_Vietnam_weighted_Protax_20180618

Questions:

1. which reserve has higher species richness (iNEXT)? 
2. nestedness
3. nmds
4. redlisted species?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)
library(vegan)
library(beanplot)
library(iNEXT)
library(RColorBrewer)
library(betapart)
sessionInfo()
```


```{r read data}
lowreadthreshold <- 100
inputfile <- c("allWWFWCS_BFCusearchMF_CROP98_otu_TAX20161217_20170518.xlsx")
leechidna <- read_excel(inputfile, sheet = "16Svert", skip = 1)
leechidna <- leechidna[ , 2:720] # can't use dplyr::select because some of the columns have duplicate names, and select can't handle that

leechidna <- leechidna %>% filter(Total_reads >= lowreadthreshold | Family %in% c("Manidae", "Lorisidae")) # filter out low-read OTUs but keep slow loris and pangolin. threshold chosen after inspection

reserve <- read_excel(inputfile, sheet = "Site_codes")
names(reserve)
reserve <- reserve %>% dplyr::select(Sample_16SVert_Row2, Worksheet_filename, Country, Protected_Area_Std_Name, Date_collected, Forest_Compartment, Original_ID_Code, Approx_number_leeches, WWF_SWG_ID)

# recode reserve names
reserve <- reserve %>% mutate(
	reservecode = case_when(
		Protected_Area_Std_Name == "BachMaExt" ~ "BME",
		Protected_Area_Std_Name == "BachMaNP" ~ "BMNP",
		Protected_Area_Std_Name == "HueSaoLaNR" ~ "HSL",
		Protected_Area_Std_Name == "Laving Lavern" ~ "LL",
		Protected_Area_Std_Name == "Phou Si Thone" ~ "PST",
		Protected_Area_Std_Name == "QuangNamNR" ~ "QNNR",
		Protected_Area_Std_Name == "QuangNamSaoLaNR" ~ "QNSL",
		Protected_Area_Std_Name == "XeSap" ~ "XS"
	) %>% as.factor()
)

levels(reserve$reservecode)

# combining reserves
reserve <- reserve %>% mutate(
	reservecode2 = case_when(
		Protected_Area_Std_Name == "BachMaExt" ~ "BM",
		Protected_Area_Std_Name == "BachMaNP" ~ "BM",
		Protected_Area_Std_Name == "HueSaoLaNR" ~ "HSL",
		Protected_Area_Std_Name == "Laving Lavern" ~ "LAO",
		Protected_Area_Std_Name == "Phou Si Thone" ~ "LAO",
		Protected_Area_Std_Name == "QuangNamNR" ~ "QNSL",
		Protected_Area_Std_Name == "QuangNamSaoLaNR" ~ "QNSL",
		Protected_Area_Std_Name == "XeSap" ~ "LAO"
	) %>% as.factor()
)

levels(reserve$reservecode2)

# previous code using 20170305 version
# inputfile <- c("allWWFWCS_BFCusearchMF_CROP98_otu_TAX20161217_20170305.xlsx")
# reserve <- reserve %>% dplyr::select(Sample_16SVert_Row2, Reserve_StdName, Reserve_SubName, Approx_number_leeches)
```


```{r sum read numbers by taxonomy}
names(leechidna)
leechidna2 <- leechidna %>% group_by(OTUs_merged_by_taxonomy) %>% summarise_at(vars(WWF1:WWF465), sum)
names(leechidna2)
# leechidna2$OTUs_merged_by_taxonomy
```


```{r remove negative controls}
negctrls <- c("BM","H1s","H2s","H3s","Q1s","Q2s","Q3s","XS2","XS5","XS6","XS7","XS8","XS9","SWG1s","SWG2s","XS1s","XS2s","2015Dec")

leechidna3 <- leechidna2

for(i in 1:length(negctrls)) {
	leechidna3 <- leechidna3 %>% dplyr::select(-starts_with(negctrls[i]))
}

# the slow way
# names(leechidna3)
# leechidna3 <- leechidna2 %>% select(-starts_with(neg[1]), -starts_with(neg[2]), -starts_with(neg[3]), -starts_with(neg[4]), -starts_with(neg[5]), -starts_with(neg[6]), -starts_with(neg[7]), -starts_with(neg[8]), -starts_with(neg[9]), -starts_with(neg[10]), -starts_with(neg[11]), -starts_with(neg[12]), -starts_with(neg[13]), -starts_with(neg[14]), -starts_with(neg[15]), -starts_with(neg[16]), -starts_with(neg[17]), -starts_with(neg[18])) 
```


```{r filter by read number, eval=FALSE}
# if i want to set cells to 0 that have fewer reads than the negative controls for that batch of samples.  I haven't implemented it yet, but i have generated the negative control matrix.
leechidna_negctrls <- leechidna2
leechidna_samples <- colnames(leechidna3)
leechidna_negctrls <- leechidna_negctrls %>% dplyr::select(-one_of(leechidna_samples))
# leechidna_negctrls
```


```{r add total_reads and incidence columns}
leechidna3$total_reads <- leechidna3 %>% dplyr::select(-OTUs_merged_by_taxonomy) %>% rowSums()

leechidna3$incidence <- leechidna3 %>% dplyr::select(-OTUs_merged_by_taxonomy, -total_reads) %>% specnumber(MARGIN=1)

names(leechidna3)
# leechidna3$incidence
# leechidna3$total_reads
```


```{r filter by taxonomy}
# change OTU names to format that is acceptable to R as column names
leechidna3$OTUs_merged_by_taxonomy <- make.names(leechidna3$OTUs_merged_by_taxonomy, unique = TRUE)
leechidna3$OTUs_merged_by_taxonomy <- gsub("cf..", "cf_", leechidna3$OTUs_merged_by_taxonomy, fixed = TRUE) # fixed = TRUE is to not use grep

# remove some taxa
taxa_to_remove <- c("Actinopterygii|Amphibia|Archelosauria|lab.contamination|laboratory.contamination")
leechidna4 <- leechidna3 %>% filter(!grepl(taxa_to_remove, OTUs_merged_by_taxonomy))
# leechidna4$OTUs_merged_by_taxonomy

# taxonomic subsets
leechidna_mammals <- leechidna4 %>% filter(!grepl(c("no.assignment|root"), OTUs_merged_by_taxonomy))

leechidna_carnivora <- leechidna4 %>% filter(grepl(c("Carnivora"), OTUs_merged_by_taxonomy))

leechidna_critters <- leechidna4 %>% filter(grepl(c("Rodentia|Scandentia|Lagomorpha|Insectivora|Pholidota"), OTUs_merged_by_taxonomy))

leechidna_artiodactyla <- leechidna4 %>% filter(grepl(c("Artiodactyla"), OTUs_merged_by_taxonomy))

leechidna_primates <- leechidna4 %>% filter(grepl(c("Primates"), OTUs_merged_by_taxonomy)) 
names(leechidna_primates)
```


```{r names of taxonomic subsets}
varNames <- c("mammals", "carnivora", "critters", "artiodactyla", "primates")
```


```{r transpose to community}
# create site X species tables from the species X site tables
# filename outputs:  leechidna_carnivora -> leechidna_carnivora_t  # site X OTU tables

for (i in seq_along(varNames)) {
  d <- get(paste0("leechidna_", varNames[i])) # get a community subset
  d_t <- t(d) # transpose
  colnames(d_t) <- d_t[1,] # move taxonomic names to column names
  d_t <- as.data.frame(d_t) # coerce to dataframe
  d_t <- rownames_to_column(d_t, var = "site") # add site names to dataframe.  site = "WWF1"
  d_t <- d_t %>% filter(site != "OTUs_merged_by_taxonomy", site != "incidence", site != "total_reads") # filter out these rows
  assign(paste0("leechidna_", varNames[i], "_t"), d_t) # assign output to the correct object name
}
```

https://codedocean.wordpress.com/2014/01/14/r-working-with-named-objects-in-a-loop/

```{r create environment and Site_by_OTU datasets}
# filename outputs:  Site_by_OTU_carnivora and reserve_carnivora
for (i in seq_along(varNames)) {
  d <- get(paste0("leechidna_", varNames[i], "_t")) # get a community subset
  dreserve <- left_join(d, reserve, by = c("site" = "Sample_16SVert_Row2")) # use site (e.g. "WWF1") to append environment data in correct order
  assign(paste0("Site_by_OTU_", varNames[i]), dreserve) # create site X OTU tables for outside use (includes env data)
  dreserve <- dreserve %>% select(site, Protected_Area_Std_Name, reservecode, reservecode2, Date_collected, Forest_Compartment, Original_ID_Code, Approx_number_leeches, WWF_SWG_ID) # choose the env columns
  # previous line
	# dreserve <- dreserve %>% select(site, Reserve_StdName, Approx_number_leeches) # choose the env columns
  assign(paste0("reserve_", varNames[i]), dreserve) # create env dataset with correct name
  d <- d %>% select(-site)
  assign(paste0("leechidna_", varNames[i], "_t"), d)
}

# now that site column has been removed from leechidna_carnivora_t, convert the remaining OTU columns to numeric
# http://stackoverflow.com/questions/2288485/how-to-convert-a-data-frame-column-to-numeric-type
for (i in seq_along(varNames)) {
  d <- get(paste0("leechidna_", varNames[i], "_t")) # get a community subset
  d <- as.data.frame(sapply(d, function(x) as.numeric(as.character(x)))) 
  assign(paste0("leechidna_", varNames[i], "_t"), d)
}

# omit rows that have rowSums = 0, from both OTU and environment datsets
for (i in seq_along(varNames)) {
  d <- get(paste0("leechidna_", varNames[i], "_t")) # get a community subset
  r <- get(paste0("reserve_", varNames[i]))
  r <- r[which(rowSums(d) > 0), ]
  d <- d[which(rowSums(d) > 0), ]
  assign(paste0("leechidna_", varNames[i], "_t"), d)
  assign(paste0("reserve_", varNames[i]), r)
}

rm(otu)
rm(d)
rm(r)
rm(d_t)
rm(dreserve)
```

```{r Site_by_OTU tables}
# change column orders of the site X OTU tables and write to disk
for (i in seq_along(varNames)) {
  otu <- get(paste0("Site_by_OTU_", varNames[i])) # get a community subset
  otu <- otu %>% dplyr::select(site, Worksheet_filename, Country, reservecode, reservecode2, Protected_Area_Std_Name, Date_collected, Forest_Compartment, Original_ID_Code, Approx_number_leeches, WWF_SWG_ID, everything()) # change the column orders
  otu <- otu %>% rename(ECEC_code = site)
  assign(paste0("Site_by_OTU_", varNames[i]), otu) # create object with correct name
  write.table(get(paste0("Site_by_OTU_", varNames[i])), file = paste0("Site_by_OTU_", varNames[i],".tsv"),sep ="\t", row.names = FALSE)
}
```


### Analyses

```{r, eval = FALSE}
xtabs(~ reservecode, reserve_mammals)
 # BME BMNP  HSL   LL  PST QNNR QNSL   XS 
 #  12   11  151   28   66   19  161   38 

xtabs(~ reservecode2, reserve_mammals)
#   BM  HSL  LAO QNSL 
#   23  151  132  180 
```

```{r make binary communities}
for (i in seq_along(varNames)) {
  otu <- get(paste0("leechidna_", varNames[i], "_t")) # get a community subset
  otu[otu > 1] <- 1 # convert to presence/absence
  assign(paste0("leechidna_", varNames[i], "_tb"), otu) # create object with correct name
}
```

#### specpool
```{r specpool}
taxon <- "mammals" 
# c("mammals", "carnivora", "critters", "artiodactyla", "primates")

# choose taxon subdataset for downstream analyses
leechidna_tb <- get(str_c("leechidna_", taxon, "_tb"))
reserve_tb <- get(str_c("reserve_", taxon))

specpool(leechidna_tb, reserve_tb$reservecode2)
specpool(leechidna_tb, reserve_tb$reservecode)

par(mfrow=c(2,1))
beanplot(specnumber(leechidna_tb) ~ reserve_tb$reservecode, col = c("grey", "white"))
beanplot(specnumber(leechidna_tb) ~ reserve_tb$reservecode2, col = c("grey", "white"))
par(mfrow=c(1,1))
```

Mammals
Species     chao  chao.se    jack1  jack1.se    jack2     boot   boot.se   n
BME       49 51.24583 2.373260 55.41667 3.8792539 53.67424 52.81269 2.5403089  12
BMNP      46 53.42424 5.269337 58.72727 6.3766102 61.34545 52.41017 4.4653216  11
HSL       60 60.66225 1.298306 61.98675 1.4048479 61.01974 61.26154 0.9676567 151
LL        49 53.88170 4.295591 57.67857 4.0103501 58.88228 53.64013 2.9011469  28
PST       60 60.63312 1.106720 62.95455 1.7058076 59.18019 62.26494 1.4489257  66
QNNR      55 61.06316 5.638836 62.57895 4.0808994 65.51170 58.61652 2.4183328  19
QNSL      63 63.12422 0.435669 63.99379 0.9937888 61.05575 63.98016 0.9765294 161
XS        53 57.92928 4.331917 61.76316 3.5252355 62.91536 57.58724 2.6133584  38

     Species     chao   chao.se    jack1 jack1.se    jack2     boot   boot.se   n
BM        58 75.21739 13.789122 69.47826 4.738931 76.94862 63.15751 2.7936120  23
HSL       60 60.66225  1.298306 61.98675 1.404848 61.01974 61.26154 0.9676567 151
LAO       69 75.07860  6.035715 75.94697 2.625708 78.93159 72.26380 1.5540458 132
QNSL      63 63.00000  0.000000 63.00000 0.000000 59.06654 63.62080 0.7846341 180


#### Species richness by Reserve:  iNEXT analysis
input is the Site X OTU table (e.g. leechidna_mammals_t), which is subset into habitats by the reserve_named_mammals$reservecode or reservecode2 
```{r inext fxn reservecode, eval = FALSE}
inextcomm <- function(taxon = "mammals") { 
	# taxon == c("mammals", "carnivora", "critters", "artiodactyla", "primates")

	# choose taxon subdataset for downstream analyses
	leechidna_tb <- get(str_c("leechidna_", taxon, "_tb"))
	reserve_tb <- get(str_c("reserve_", taxon))

	# reserve subsets
	BME <- leechidna_tb %>% dplyr::filter(reserve_tb$reservecode %in% c("BME"))
	BMNP <- leechidna_tb %>% dplyr::filter(reserve_tb$reservecode %in% c("BMNP"))
	HSL <- leechidna_tb %>% dplyr::filter(reserve_tb$reservecode %in% c("HSL"))
	LL <- leechidna_tb %>% dplyr::filter(reserve_tb$reservecode %in% c("LL"))
	PST <- leechidna_tb %>% dplyr::filter(reserve_tb$reservecode %in% c("PST"))
	QNNR <- leechidna_tb %>% dplyr::filter(reserve_tb$reservecode %in% c("QNNR"))
	QNSL <- leechidna_tb %>% dplyr::filter(reserve_tb$reservecode %in% c("QNSL"))
	XS <- leechidna_tb %>% dplyr::filter(reserve_tb$reservecode %in% c("XS"))

	cname <- c("BME", "BMNP", "HSL", "LL", "PST", "QNNR", "QNSL", "XS")

	comm4inext <- matrix(c(colSums(BME), colSums(BMNP), colSums(HSL), colSums(LL), colSums(PST), colSums(QNNR), colSums(QNSL), colSums(XS)), ncol = 8)
	# rownames(comm4inext) <- rname
	colnames(comm4inext) <- cname
	comm4inext <- rbind(c(nrow(BME), nrow(BMNP), nrow(HSL), nrow(LL), nrow(PST), nrow(QNNR), nrow(QNSL), nrow(XS)), comm4inext)
	comm4inext
}

comm4inext <- inextcomm("artiodactyla")
comm4inext
	# taxondb == c("named_mammals", "carnivora", "critters", "artiodactyla", "primates")
confnum=0.95 # set confidence here

outcomm0 <- iNEXT(comm4inext, q=0, conf=confnum, datatype="incidence_freq")
# Hill numbers:  0 = sp richness, 1 = Shannon, 2 = inverse Simpson
outcomm0$DataInfo
ChaoRichness(comm4inext, datatype="incidence_freq") # same as specpool results, so i trust that we have done this correctly
ChaoShannon(comm4inext, datatype="incidence_freq")

outcomm1 <- iNEXT(comm4inext, q=1, conf=confnum, datatype="incidence_freq")
ggiNEXT(outcomm1, type=1)
outcomm2 <- iNEXT(comm4inext, q=2, conf=confnum, datatype="incidence_freq")
ggiNEXT(outcomm2, type=1)

ggiNEXT(outcomm0, type=1) # sample-size-based rarefaction/extrapolation curve:  species richness
ggiNEXT(outcomm1, type=1) # sample-size-based rarefaction/extrapolation curve:  Shannon

ggiNEXT(outcomm0, type=2) # sample completeness curve
ggiNEXT(outcomm0, type=3) # coverage-based rarefaction/extrapolation curve
# outcomm0$DataInfo
# outcomm0$iNextEst
# outcomm0$AsyEst
```

```{r inext fxn reservecode2, eval = FALSE}
inextcomm <- function(taxon = "mammals") { 
	# taxon == c("mammals", "carnivora", "critters", "artiodactyla", "primates")

	# choose taxon subdataset for downstream analyses
	leechidna_tb <- get(str_c("leechidna_", taxon, "_tb"))
	reserve_tb <- get(str_c("reserve_", taxon))

	# reserve subsets
	BM <- leechidna_tb %>% dplyr::filter(reserve_tb$reservecode2 %in% c("BM"))
	HSL <- leechidna_tb %>% dplyr::filter(reserve_tb$reservecode2 %in% c("HSL"))
	LAO <- leechidna_tb %>% dplyr::filter(reserve_tb$reservecode2 %in% c("LAO"))
	QNSL <- leechidna_tb %>% dplyr::filter(reserve_tb$reservecode2 %in% c("QNSL"))

	cname <- c("BM", "HSL", "LAO", "QNSL")

	comm4inext <- matrix(c(colSums(BM), colSums(HSL), colSums(LAO), colSums(QNSL)), ncol = 4)
	colnames(comm4inext) <- cname
	comm4inext <- rbind(c(nrow(BM), nrow(HSL), nrow(LAO), nrow(QNSL)), comm4inext)
	comm4inext
}


comm4inext <- inextcomm("mammals")
comm4inext
	# taxon == c("mammals", "carnivora", "critters", "artiodactyla", "primates")
confnum=0.95 # set confidence here

outcomm0 <- iNEXT(comm4inext, q=0, conf=confnum, datatype="incidence_freq")
# Hill numbers:  0 = sp richness, 1 = Shannon, 2 = inverse Simpson
outcomm0$DataInfo
ChaoRichness(comm4inext, datatype="incidence_freq") # same as specpool results, so i trust that we have done this correctly
ChaoShannon(comm4inext, datatype="incidence_freq")

outcomm1 <- iNEXT(comm4inext, q=1, conf=confnum, datatype="incidence_freq")
ggiNEXT(outcomm1, type=1)
outcomm2 <- iNEXT(comm4inext, q=2, conf=confnum, datatype="incidence_freq")
ggiNEXT(outcomm2, type=1)

ggiNEXT(outcomm0, type=1) # sample-size-based rarefaction/extrapolation curve:  species richness
ggiNEXT(outcomm1, type=1) # sample-size-based rarefaction/extrapolation curve:  Shannon

ggiNEXT(outcomm0, type=2) # sample completeness curve
ggiNEXT(outcomm0, type=3) # coverage-based rarefaction/extrapolation curve
# outcomm0$DataInfo
# outcomm0$iNextEst
# outcomm0$AsyEst
```


#### Community analyses

#### Run tabasco before removing zerotons and singletons
```{r}
taxon <- "mammals"
leechidna_tb <- get(str_c("leechidna_", taxon, "_tb"))
reserve_tb <- get(str_c("reserve_", taxon))

community.jmds <- metaMDS(leechidna_tb, distance = "jaccard", trymax = 40, binary=TRUE)
community.jmds <- metaMDS(leechidna_tb, distance = "jaccard", binary = TRUE, previous.best = community.jmds)  
stressplot(community.jmds)
tabasco(leechidna_tb, use = community.jmds, labCol = reserve_tb$reservecode2)
```

No visual sign of nestedness, but the betapart analysis finds nestedness

#### Run betapart analyses before removing zerotons and singletons
```{r}
taxon <- "mammals"
leechidna_tb <- get(str_c("leechidna_", taxon, "_tb"))
reserve_tb <- get(str_c("reserve_", taxon))

# no nestedness 
(ALL.dist <- leechidna_tb %>% beta.multi(index.family="jac"))
# $beta.JTU
# [1] 0.9941085
# $beta.JNE
# [1] 0.002525642
# $beta.JAC
# [1] 0.9966341


ALL.dist <- leechidna_tb %>% beta.pair(index.family="jac")
# turnover only
ALL.dist.subset <- ALL.dist[["beta.jtu"]]
ALL.dist.jtu.jmds <- metaMDS(ALL.dist.subset)
ALL.dist.jtu.jmds <- metaMDS(ALL.dist.subset, previous.best = ALL.dist.jtu.jmds)
stressplot(ALL.dist.jtu.jmds)
# nestedness only
ALL.dist.subset <- ALL.dist[["beta.sne"]]
ALL.dist.jne.jmds <- metaMDS(ALL.dist.subset)
ALL.dist.jne.jmds <- metaMDS(ALL.dist.subset, previous.best = ALL.dist.jne.jmds)
stressplot(ALL.dist.jne.jmds)

par(mfrow=c(2,2))
plot(community.jmds, main = "All beta diversity", ylim = c(-1.0, 1.0))

plot(ALL.dist.sim.jmds, main = "Turnover beta diversity only", ylim = c(-1.0, 1.0))

plot(ALL.dist.sne.jmds, main = "Nestedness beta diversity only", ylim = c(-1.0, 1.0))
par(mfrow=c(1,1))
```


#### Unconstrained ordination
#### For NMDS, mvabund, and boral analyses, remove zerotons and singletons
Should be left with 358 species (variables)
```{r remove zerotons and singletons from communityB}
taxon <- "mammals"
leechidna_tb <- get(str_c("leechidna_", taxon, "_tb"))
reserve_tb <- get(str_c("reserve_", taxon))

leechidna_tb <- leechidna_tb[, which(specnumber(leechidna_tb, MARGIN=2) > 1)]
```

## NMDS ordination
```{r NMDS}
### do NMDS analysis to quickly see patterns ####
community.jmds <- metaMDS(leechidna_tb, distance = "jaccard", trymax = 40, binary=TRUE)
community.jmds <- metaMDS(leechidna_tb, distance = "jaccard", binary = TRUE, previous.best = community.jmds)  # doesn't converge well, with final stress > 0.20
stressplot(community.jmds)
```


```{r ordispider plot reservecode}
#### plot the communities 
levels(reserve_tb$reservecode)
# BME BMNP HSL LL  PST QNNR QNSL   XS 
levels(reserve_tb$reservecode2)
# BM  HSL  LAO QNSL 

levels(reserve_tb$reservecode)
(sprichness <- specnumber(leechidna_tb, MARGIN = 1)) # number of species per site
colvec <- brewer.pal(5, "Set1")

plot(community.jmds, display="sites", type="n")

with(reserve_tb, ordispider(community.jmds, reservecode, cex=.5, draw="polygon", col=as.character(colvec[1]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("BME"))))

with(reserve_tb, ordispider(community.jmds, reservecode, cex=.5, draw="polygon", col=as.character(colvec[1]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("BMNP"))))

with(reserve_tb, ordispider(community.jmds, reservecode, cex=.5, draw="polygon", col=as.character(colvec[2]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("HSL"))))

with(reserve_tb, ordispider(community.jmds, reservecode, cex=.5, draw="polygon", col=as.character(colvec[3]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("LL"))))

with(reserve_tb, ordispider(community.jmds, reservecode, cex=.5, draw="polygon", col=as.character(colvec[3]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("PST"))))

with(reserve_tb, ordispider(community.jmds, reservecode, cex=.5, draw="polygon", col=as.character(colvec[3]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("XS"))))

with(reserve_tb, ordispider(community.jmds, reservecode, cex=.5, draw="polygon", col=as.character(colvec[4]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("QNNR"))))

with(reserve_tb, ordispider(community.jmds, reservecode, cex=.5, draw="polygon", col=as.character(colvec[4]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("QNSL"))))
```


```{r ordispider plot reservecode2}
#### plot the communities 
levels(reserve_tb$reservecode)
# BME BMNP HSL LL  PST QNNR QNSL   XS 
levels(reserve_tb$reservecode2)
# BM  HSL  LAO QNSL 

levels(reserve_tb$reservecode)
(sprichness <- specnumber(leechidna_tb, MARGIN = 1)) # number of species per site
colvec <- brewer.pal(5, "Set1")

plot(community.jmds, display="sites", type="n")

with(reserve_tb, ordispider(community.jmds, reservecode2, cex=.5, draw="polygon", col=as.character(colvec[1]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("BM"))))

with(reserve_tb, ordispider(community.jmds, reservecode2, cex=.5, draw="polygon", col=as.character(colvec[2]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("HSL"))))

with(reserve_tb, ordispider(community.jmds, reservecode2, cex=.5, draw="polygon", col=as.character(colvec[3]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("LAO"))))

with(reserve_tb, ordispider(community.jmds, reservecode2, cex=.5, draw="polygon", col=as.character(colvec[4]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("QNSL"))))

```





