---
title: "Vietnam Leeches using Alex Crampton-Platt's 20180618 Protax run"
author: "Douglas Yu"
date: "`r Sys.Date()`"
output: html_document
---

This protax run was run using the pipeline by Axtner et al., which starts with the Midori database and set high priors for Vietnam species.  

The data files are in the folder:  Vietnam_leeches_swarm_lulu_20180404/
OTU table:  2015WWFWCS_swarm_lulu_otu_table_20180404.xlsx
OTU representative seqs:  all_2015WWFWCS_otu_table_swarm_lulu_vert16S_20180404.fas
Sample-to-Site codings:  allWWFWCS_sample_codes_20180621.xlsx
Protax output files are in:  Vietnam_leeches_swarm_lulu_20180404/doug_vietnam_16S_Vietnam_weighted_Protax_20180618/

Analysis output is in:  analysis/

Questions:

1. which reserve has higher species richness and diversity (iNEXT)? 
2. nestedness or turnover?
3. nmds
4. redlisted species?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)
library(vegan)
library(beanplot)
library(iNEXT)
library(RColorBrewer)
library(betapart)
library(cowplot)
sessionInfo()
```

```{r data folders}
folder1 <- "Vietnam_leeches_swarm_lulu_20180404"
folder2 <- "doug_vietnam_16S_Vietnam_weighted_Protax_20180618"
```

```{r read sample to site database and add reserve codes} 
reserve <- read_excel(file.path(folder1, "allWWFWCS_sample_codes_20180621.xlsx"), sheet = "Site_codes")
names(reserve)
reserve <- reserve %>% dplyr::select(Sample_16SVert_Row2, Worksheet_filename, Country, Protected_Area_Std_Name, Date_collected, Forest_Compartment, Original_ID_Code, Approx_number_leeches, WWF_SWG_ID)

# recode reserve names to reservecode
reserve <- reserve %>% mutate(
	reservecode = case_when(
		Protected_Area_Std_Name == "BachMaExt" ~ "BM",
		Protected_Area_Std_Name == "BachMaNP" ~ "BM",
		Protected_Area_Std_Name == "HueSaoLaNR" ~ "HSL",
		Protected_Area_Std_Name == "Hue SNR" ~ "HSL",
		Protected_Area_Std_Name == "Quang Nam SNR" ~ "QNSL",
		Protected_Area_Std_Name == "Laving Lavern" ~ "LL",
		Protected_Area_Std_Name == "Phou Si Thone" ~ "PST",
		Protected_Area_Std_Name == "QuangNamNR" ~ "QNSL",
		Protected_Area_Std_Name == "QuangNamSaoLaNR" ~ "QNSL",
		Protected_Area_Std_Name == "XeSap" ~ "XS"
	) %>% as.factor()
)

levels(reserve$reservecode)

# recode reserve names to reservecode2
reserve <- reserve %>% mutate(
	reservecode2 = case_when(
		Protected_Area_Std_Name == "BachMaExt" ~ "BME",
		Protected_Area_Std_Name == "BachMaNP" ~ "BMNP",
		Protected_Area_Std_Name == "HueSaoLaNR" ~ "HSL",
		Protected_Area_Std_Name == "Hue SNR" ~ "HSL",
		Protected_Area_Std_Name == "Quang Nam SNR" ~ "QNSL",
		Protected_Area_Std_Name == "Laving Lavern" ~ "LL",
		Protected_Area_Std_Name == "Phou Si Thone" ~ "PST",
		Protected_Area_Std_Name == "QuangNamNR" ~ "QNSL",
		Protected_Area_Std_Name == "QuangNamSaoLaNR" ~ "QNSL",
		Protected_Area_Std_Name == "XeSap" ~ "XS"
	) %>% as.factor()
)

levels(reserve$reservecode2)

# recode reserve names to reservecode3
reserve <- reserve %>% mutate(
	reservecode3 = case_when(
		Protected_Area_Std_Name == "BachMaExt" ~ "BMHQNXS",
		Protected_Area_Std_Name == "BachMaNP" ~ "BMHQNXS",
		Protected_Area_Std_Name == "HueSaoLaNR" ~ "BMHQNXS",
		Protected_Area_Std_Name == "Hue SNR" ~ "BMHQNXS",
		Protected_Area_Std_Name == "Quang Nam SNR" ~ "BMHQNXS",
		Protected_Area_Std_Name == "Laving Lavern" ~ "LL",
		Protected_Area_Std_Name == "Phou Si Thone" ~ "PST",
		Protected_Area_Std_Name == "QuangNamNR" ~ "BMHQNXS",
		Protected_Area_Std_Name == "QuangNamSaoLaNR" ~ "BMHQNXS",
		Protected_Area_Std_Name == "XeSap" ~ "BMHQNXS"
	) %>% as.factor()
)

```

Now, skip down to code chunk:  "1 write leechotu file after visual check" before 

```{r read in protax tables, eval = FALSE}
rank <- c("order", "family", "genus", "species")

loadFile1 <- function(rank) { 
   filename <- str_c("all_2015WWFWCS_otu_table_swarm_lulu_vert16S_20180404.fas.w_", rank, "_probs") 
   # e.g. "all_2015WWFWCS_otu_table_swarm_lulu_vert16S_20180404.fas.w_order_probs"

   df <- read_tsv(file.path(folder1, folder2, filename), 
   	col_names = c("OTU", "taxid", "logprob", "rank", "protax"), 
  	na = "NA",
	  col_types = cols(
	     OTU = col_character(),
	     taxid = col_number(),
	     logprob = col_double(),
	     rank = col_integer(),
	     protax = col_character()
	     )
	 	)

    df$prob <- exp(df$logprob) # add prob column
    
    df <- df %>% dplyr::select(OTU, taxid, logprob, prob, rank, protax)

    df # output is df
}

for(rank in rank) {
    assign(str_c("protax_", rank), loadFile1(rank))
}
```

```{r filter and join protax tables, eval = FALSE}
minprob <- 0
protax_order <- protax_order %>% dplyr::filter(prob >= minprob)
protax_family <- protax_family %>% dplyr::filter(prob >= minprob)
protax_genus <- protax_genus %>% dplyr::filter(prob >= minprob)
protax_species <- protax_species %>% dplyr::filter(prob >= minprob)

protax_order_family <- left_join(protax_order, protax_family, by = c("OTU"), suffix = c(".order", ".family"))

protax_genus_species <- left_join(protax_genus, protax_species, by = c("OTU"), suffix = c(".genus", ".species"))

protax_all <- left_join(protax_order_family, protax_genus_species, by = c("OTU"))

rm(protax_order_family, protax_genus_species, protax_order, protax_family, protax_genus, protax_species)

# names(protax_all)

# delete rows with NA in the protax.order, protax.family, protax.genus
protax_all <- protax_all %>% drop_na(protax.order, protax.family, protax.genus)

# copy values of protax.species into protax.consensus if protax.species is not NA, otherwise, just copy value of protax.genus into protax.consensus
protax_all <- protax_all %>% mutate(protax.consensus = ifelse(!is.na(protax.species), protax.species, protax.genus))

# conditionally copy the probability value
# protax_all <- protax_all %>% mutate(prob.consensus = ifelse(!is.na(prob.species), prob.species, prob.genus))

# create column with source of consensus taxonomy
protax_all <- protax_all %>% mutate(source.consensus = ifelse(!is.na(prob.species), "species", "genus"))

protax_all <- protax_all %>% dplyr::select(OTU, protax.consensus, prob.order, prob.family, prob.genus, prob.species, source.consensus)

names(protax_all)

```

```{r read otu table and merge with protax_all, eval = FALSE}
inputfile <- c("2015WWFWCS_swarm_lulu_otu_table_20180404.xlsx")
leechotu <- read_excel(file.path(folder1, inputfile), sheet = "filtered_otus")
# names(leechotu)
leechotu <- left_join(protax_all, leechotu, by = "OTU")
leechotu <- leechotu %>% dplyr::select(-(pct_identity:OTU_2))
# names(leechotu)
total_reads <- leechotu %>% dplyr::select(BM10x1:XS9x8) %>% rowSums() %>% data_frame()
colnames(total_reads) <- "total_reads"
incidence <- leechotu %>% dplyr::select(BM10x1:XS9x8) %>% specnumber(MARGIN=1) %>% data_frame()
colnames(incidence) <- "incidence"
leechotu <- bind_cols(total_reads, incidence, leechotu)

names(leechotu)
rm(incidence)
rm(total_reads)
rm(protax_all)
```

1. now write leechotu file to disk, and switch to Excel, where I visually inspect the protax assignments and the distribution of reads over samples and manually decide final.taxonomy  
2. sum read numbers by taxonomy and remove cell values <= 2 reads
3. save negative control samples
4. remove negative control samples from leechotu4
5. filter by taxonomy and remove samples with 0 or 1 species only

```{r 1 write leechotu file after visual check}
# write_tsv(leechotu, "analysis/leechotu_protax.tsv")

leechotu2 <- read_excel("analysis/leechotu_protax.xlsx", sheet = "leechotu")
names(leechotu2)
  #  [1] "total_reads"      "incidence"        "OTU"              "protax.consensus" "prob.order"      
  #  [6] "prob.family"      "prob.genus"       "prob.species"     "source.consensus" "final.taxonomy"  
  # [11] "notes"            "RDP_MIDORI_REF"   "BM10x1"           "BM10x10"          "BM10x3"          
  # [16] "BM10x7"           "BM1x1"            "BM1x2"            "BM1x3"            "BM1x4"    
```

```{r 2 sum read numbers by taxonomy}
leechotu3 <- leechotu2 %>% 
	group_by(final.taxonomy) %>% 
	summarise_at(vars(BM10x1:XS9x8), sum) %>%  # BM10x1:XS9x8 is all the sample codes
	dplyr::filter(final.taxonomy != "OMIT")
names(leechotu3)

# change OTU names to format that is acceptable to R as column names
leechotu3$final.taxonomy <- make.names(leechotu3$final.taxonomy, unique = TRUE)

# set small cell values (1 or 2) to 0.  this should reduce incidence of species
leechotu3 <- leechotu3 %>% column_to_rownames(var = "final.taxonomy")
leechotu3[leechotu3 <= 2] <- 0
leechotu3 <- rownames_to_column(leechotu3, var = "final.taxonomy")
```

```{r 3 save negative control samples}
negctrls <- c("X2015Dec","BM","H1s","H2s","H3s","Q1s","Q2s","Q3s","SWG1s","SWG2s","XS1s","XS2s","XS2x","XS5x","XS6x","XS7x", "XS8x", "XS9x", "XS10x", "XS1x", "XS3x") # "XS10x", "XS1x", "XS3x" inferred as negctrls from having no reads in them and from the column names:  should be 16 more removed samples from these three headings

# save negative controls into leechotu_negctrls (loop doesn't work. i think because this version is an OR, whereas a loop is an AND?)
leechotu_negctrls <- leechotu3 %>% select(final.taxonomy, starts_with(negctrls[1]), starts_with(negctrls[2]), starts_with(negctrls[3]), starts_with(negctrls[4]), starts_with(negctrls[5]), starts_with(negctrls[6]), starts_with(negctrls[7]), starts_with(negctrls[8]), starts_with(negctrls[9]), starts_with(negctrls[10]), starts_with(negctrls[11]), starts_with(negctrls[12]), starts_with(negctrls[13]), starts_with(negctrls[14]), starts_with(negctrls[15]), starts_with(negctrls[16]), starts_with(negctrls[17]), starts_with(negctrls[18]), starts_with(negctrls[19]), starts_with(negctrls[20]), starts_with(negctrls[21]))

# add some columns
leechotu_negctrls$total_reads <- leechotu_negctrls %>% select(-final.taxonomy) %>% rowSums() 
leechotu_negctrls$incidence <- leechotu_negctrls %>% dplyr::select(-final.taxonomy, -total_reads) %>% specnumber(MARGIN=1)
# add a column with the max number of reads per OTU. this is a pain
otusonly <- leechotu_negctrls %>% dplyr::select(-final.taxonomy, -total_reads, -incidence) # make an OTUonly table
maxotu <- apply(X = otusonly, MARGIN = 1, FUN = max) %>% data_frame() # use apply() to apply max() to each row
colnames(maxotu) <- "maxotu" # add colname
leechotu_negctrls <- bind_cols(leechotu_negctrls, maxotu) # add back to dataframe

leechotu_negctrls <- leechotu_negctrls %>% dplyr::select(final.taxonomy, total_reads, incidence, maxotu,  everything()) # reorder columns
names(leechotu_negctrls)
rm(otusonly, maxotu)
# write_tsv(leechotu_negctrls, "analysis/leechotu_negctrls.tsv")
```

```{r 4 remove negative control samples from leechotu4}
leechotu4 <- leechotu3
# cycles through the dataset, each time removing columns with a new prefix, until all negctrl prefixes have been removed
for(i in 1:length(negctrls)) {
	leechotu4 <- leechotu4 %>% dplyr::select(-starts_with(negctrls[i])) 
}

# (551-4=) 547 negative ctrl samples + (543-1 =) 542 real samples = 1089 total samples. leechotu3 is 1089 sample columns + 1 final.taxonomy column = 1090 columns.  correct

#  and add some columns
leechotu4$total_reads <- leechotu4 %>% dplyr::select(-final.taxonomy) %>% rowSums()
leechotu4$incidence <- leechotu4 %>% dplyr::select(-final.taxonomy, -total_reads) %>% specnumber(MARGIN=1)

# add a column with the max number of reads per OTU. this is a pain
otusonly <- leechotu4 %>% dplyr::select(-final.taxonomy, -total_reads, -incidence) # make an OTUonly table
maxotu <- apply(X = otusonly, MARGIN = 1, FUN = max) %>% data_frame() # use apply() to apply max() to each row
colnames(maxotu) <- "maxotu" # add colname
leechotu4 <- bind_cols(leechotu4, maxotu) # add back to dataframe


leechotu4 <- leechotu4 %>% dplyr::select(final.taxonomy, total_reads, incidence, maxotu, everything()) # reorder columns
names(leechotu4)  # SWG1:XS9B
```

```{r filter by read number, eval=FALSE}
# this is where i could add code to set to 0 cells that have fewer reads than the max number of reads for that species in the negative controls (leech_negctrls)
# this is not an easy programming job!
```

```{r 5 filter by taxonomy and remove samples with 0 or 1 species only}
# remove non-mammal taxa
taxa_to_remove <- c("Accipitriformes|Anseriformes|Anura|Columbiformes|Galliformes|Gruiformes|Passeriformes")
leechotu5 <- leechotu4 %>% filter(!grepl(taxa_to_remove, final.taxonomy)) # 55 mammalian taxa

# remove samples (columns) that have 0 or 1 species  # Margin = 1 == rows, MARGIN = 2 == cols
leechotu5.removedsamples <- leechotu5[, which(specnumber(leechotu5, MARGIN = 2) <= 1)]  
leechotu5 <- leechotu5[, which(specnumber(leechotu5, MARGIN = 2) > 1)]  

# taxonomic subsets
leechotu5_mammals <- leechotu5
# write_tsv(leechotu5, "analysis/leechotu5_mammals.tsv")

leechotu5_lgmammals <- leechotu5 %>% filter(grepl(c("Carnivora|Artiodactyla|Primates"), final.taxonomy))

leechotu5_carnivora <- leechotu5 %>% filter(grepl(c("Carnivora"), final.taxonomy))

leechotu5_artiodactyla <- leechotu5 %>% filter(grepl(c("Artiodactyla"), final.taxonomy))

leechotu5_primates <- leechotu5 %>% filter(grepl(c("Primates"), final.taxonomy)) 

leechotu5_critters <- leechotu5 %>% filter(grepl(c("Rodentia|Scandentia|Lagomorpha|Soricomorpha|Erinaceomorpha|Chiroptera"), final.taxonomy))
```

```{r 6 names of taxonomic subsets}
varNames <- c("mammals", "lgmammals", "carnivora", "critters", "artiodactyla", "primates")
```

```{r 7 transpose OTU tables to community tables}
# create sample X OTU tables from OTU X sample tables
# filename outputs:  leechotu5_carnivora -> leechotu5_carnivora_t  # sample X OTU tables

for (i in seq_along(varNames)) {
  d <- get(paste0("leechotu5_", varNames[i])) # get a community subset
  d_t <- t(d) # transpose
  colnames(d_t) <- d_t[1,] # move taxonomic names to column names
  d_t <- data.frame(d_t) # coerce to dataframe (prev. as.data.frame())
  d_t <- rownames_to_column(d_t, var = "sample") # add sample names to dataframe.  sample is a tube = "WWF1"
  d_t <- d_t %>% filter(sample != "final.taxonomy", sample != "total_reads", sample != "incidence", sample != "maxotu") # filter out these rows
  assign(paste0("leechotu5_", varNames[i], "_t"), d_t) # assign output to the correct object name
}
```

```{r 8 create environment and sample_by_OTU datasets}
# filename outputs:  sample_by_OTU_carnivora and reserve_carnivora 
for (i in seq_along(varNames)) {
  d <- get(paste0("leechotu5_", varNames[i], "_t")) # get a community table
  dreserve <- left_join(d, reserve, by = c("sample" = "Sample_16SVert_Row2")) # use sample (e.g. "WWF1") to append environment data in correct order
  assign(paste0("Sample_by_OTU_", varNames[i]), dreserve) # create sample X OTU tables for outside use (includes env data)
  dreserve <- dreserve %>% dplyr::select(sample, Protected_Area_Std_Name, reservecode, reservecode2, reservecode3, Date_collected, Forest_Compartment, Original_ID_Code, Approx_number_leeches, WWF_SWG_ID) # choose the env columns
  assign(paste0("reserve_", varNames[i]), dreserve) # create env dataset with correct name
  d <- d %>% select(-sample)
  assign(paste0("leechotu5_", varNames[i], "_t"), d)
}

# now that sample column has been removed from leechotu5_carnivora_t, convert the remaining OTU columns to numeric
# http://stackoverflow.com/questions/2288485/how-to-convert-a-data-frame-column-to-numeric-type
for (i in seq_along(varNames)) {
  d <- get(paste0("leechotu5_", varNames[i], "_t")) # get a community subset
  d <- as.data.frame(sapply(d, function(x) as.numeric(as.character(x)))) 
  assign(paste0("leechotu5_", varNames[i], "_t"), d)
}

# omit samples that have rowSums = 0 (species that are entirely absent), from both OTU and environment datsets
for (i in seq_along(varNames)) {
  d <- get(paste0("leechotu5_", varNames[i], "_t")) # get a community subset
  r <- get(paste0("reserve_", varNames[i]))
  r <- r[which(rowSums(d) > 0), ]
  d <- d[which(rowSums(d) > 0), ]
  assign(paste0("leechotu5_", varNames[i], "_t"), d)
  assign(paste0("reserve_", varNames[i]), r)
}

rm(d)
rm(r)
rm(d_t)
rm(dreserve)
```


DON'T RUN
This saves Sample_by_OTU_taxon.tsv tables to make some tables in Excel.  
```{r sample_by_OTU tables, eval = FALSE}
# change column orders of the sample X OTU tables and write to disk
for (i in seq_along(varNames)) {
  otu <- get(paste0("Sample_by_OTU_", varNames[i])) # get a community subset
  otu <- otu %>% dplyr::select(sample, Worksheet_filename, Country, reservecode, reservecode2, reservecode3, Protected_Area_Std_Name, Date_collected, Forest_Compartment, Original_ID_Code, Approx_number_leeches, WWF_SWG_ID, everything()) # change the column orders
  otu <- otu %>% rename(KIZ_code = sample)
  assign(paste0("Sample_by_OTU_", varNames[i]), otu) # create object with correct name
  write_tsv(get(paste0("Sample_by_OTU_", varNames[i])), file.path("analysis", paste0("Sample_by_OTU_", varNames[i],".tsv")))
}
```

DON'T RUN
This creates summary tables, samples summed by reservecodes
```{r, reservecode_by_OTU tables, eval = FALSE}
reservecode_by_OTU <- Sample_by_OTU_mammals %>% dplyr::select(Country, reservecode,reservecode3, Protected_Area_Std_Name, Artiodactyla.Bovidae.Bos.taurus:Soricomorpha.Talpidae.Euroscaptor.UNK)
reservecode_by_OTU$reservecode <- as.character(reservecode_by_OTU$reservecode)
reservecode_by_OTU$reservecode3 <- as.character(reservecode_by_OTU$reservecode3)
indx <- sapply(reservecode_by_OTU, is.factor) # marks columns that are OTU taxonomic assignments
reservecode_by_OTU[indx] <- lapply(reservecode_by_OTU[indx], function(x) as.numeric(as.character(x)))
reservecode_by_OTU$reservecode <- as.factor(reservecode_by_OTU$reservecode)
reservecode_by_OTU$reservecode3 <- as.factor(reservecode_by_OTU$reservecode3)

reservecode_by_OTU_readsum <- reservecode_by_OTU %>% 
	group_by_at(vars(Country, reservecode, reservecode3)) %>% 
	summarise_at(vars(Artiodactyla.Bovidae.Bos.taurus:Soricomorpha.Talpidae.Euroscaptor.UNK), sum) 

reservecode_by_OTU_b <- reservecode_by_OTU
reservecode_by_OTU_b$Country <- as.factor(reservecode_by_OTU_b$Country)
reservecode_by_OTU_b$Protected_Area_Std_Name <- as.factor(reservecode_by_OTU_b$Protected_Area_Std_Name)
reservecode_by_OTU_b[reservecode_by_OTU_b > 0] <- 1 # works since i have set all metadata cols to factor
reservecode_by_OTU_samplecount <- reservecode_by_OTU_b %>% 
	group_by_at(vars(Country, reservecode, reservecode3)) %>% 
	summarise_at(vars(Artiodactyla.Bovidae.Bos.taurus:Soricomorpha.Talpidae.Euroscaptor.UNK), sum) 

reservecode_by_OTU_samplecount2 <- reservecode_by_OTU_b %>% 
	group_by_at(vars(Country, reservecode, reservecode3)) %>% 
	summarise(samples = n()) 

reservecode_by_OTU_samplecount <- left_join(reservecode_by_OTU_samplecount2, reservecode_by_OTU_samplecount)

rm(reservecode_by_OTU_samplecount2)
rm(reservecode_by_OTU_b)
rm(reservecode_by_OTU)

# write_tsv(reservecode_by_OTU_readsum, "analysis/reservecodes_by_OTU_readsum.tsv")
# write_tsv(reservecode_by_OTU_samplecount, "analysis/reservecodes_by_OTU_samplecount.tsv")
```

DON'T RUN
Find the outlier PST sample, which is SWG53 (Mustela kathiah). i don't need to run this because i've removed this and other 1-species-only samples in the codechunk:  "filter by taxonomy and remove samples with 0 or 1 species only"
```{r find outlier PST community, eval = FALSE}

leechotu5_mammals_pst_t <- leechotu5_mammals_t %>% dplyr::filter(reserve_mammals$reservecode == "PST" & !(reserve_mammals$sample %in% c("SWG53"))) #  
reserve_mammals_pst <- reserve_mammals %>% dplyr::filter(reservecode == "PST" & !(reserve_mammals$sample %in% c("SWG53")))

community.jmds <- metaMDS(leechotu5_mammals_pst_t, distance = "jaccard", trymax = 40, binary=FALSE)
community.jmds <- metaMDS(leechotu5_mammals_pst_t, distance = "jaccard", binary = FALSE, previous.best = community.jmds)  
stressplot(community.jmds)
tabasco(leechotu5_mammals_pst_t, use = community.jmds, labCol = reserve_mammals_pst$sample)
plot(community.jmds)
text(community.jmds, display = "sites", cex=0.7, col="blue")
text(community.jmds, display = "species", cex=0.7, col="blue")
```

https://davetang.org/muse/2018/05/15/making-a-heatmap-in-r-with-the-pheatmap-package/


### Analyses

```{r, eval = FALSE}
xtabs(~ reservecode, reserve_mammals)
  # BM  HSL   LL  PST QNSL   XS 
  # 24  160   28   63  189   74 
xtabs(~ reservecode2, reserve_mammals)
 # BME BMNP  HSL   LL  PST QNSL   XS 
 #  13   11  160   28   63  189   74 
```

```{r make binary communities}
for (i in seq_along(varNames)) {
  otu <- get(paste0("leechotu5_", varNames[i], "_t")) # get a community subset
  otu[otu > 1] <- 1 # convert to presence/absence
  assign(paste0("leechotu5_", varNames[i], "_tb"), otu) # create object with correct name
}
rm(otu)
```

#### specpool
```{r specpool}
taxon <- "lgmammals" 
# c("mammals", "lgmammals", "carnivora", "critters", "artiodactyla", "primates")

# choose taxon subdataset for downstream analyses
leechotu_tb <- get(str_c("leechotu5_", taxon, "_tb"))
reserve_tb <- get(str_c("reserve_", taxon))

specpool(leechotu_tb, reserve_tb$reservecode2)
specpool(leechotu_tb, reserve_tb$reservecode)

par(mfrow=c(2,1))
beanplot(specnumber(leechotu_tb) ~ reserve_tb$reservecode, col = c("grey", "white"))
beanplot(specnumber(leechotu_tb) ~ reserve_tb$reservecode2, col = c("grey", "white"))
par(mfrow=c(1,1))
```

Mammals
specpool(leechotu_tb, reserve_tb$reservecode2)
     Species     chao   chao.se    jack1 jack1.se    jack2     boot  boot.se   n
BME       29 43.76923 15.261042 36.38462 3.241740 41.60256 32.25520 1.743129  13
BMNP      24 38.54545 15.037144 31.27273 4.185769 36.34545 27.20364 2.323023  11
HSL       36 43.95000 11.591160 39.97500 1.987500 42.94371 37.75953 1.083133 160
LL        28 29.10204  1.577984 31.85714 2.376544 29.31217 30.43135 1.587379  28
PST       37 38.10714  1.745834 39.95238 1.704558 39.04659 38.80581 1.187327  63
QNSL      37 37.49735  1.026605 38.98942 1.406731 37.03163 38.37930 1.012102 189
XS        32 40.05631  8.171844 38.90541 2.609998 42.83728 35.13200 1.445045  74
specpool(leechotu_tb, reserve_tb$reservecode)
     Species     chao   chao.se    jack1 jack1.se    jack2     boot  boot.se   n
BM        33 39.13333  5.697875 40.66667 3.343734 43.61594 36.62165 1.842147  24
HSL       36 43.95000 11.591160 39.97500 1.987500 42.94371 37.75953 1.083133 160
LL        28 29.10204  1.577984 31.85714 2.376544 29.31217 30.43135 1.587379  28
PST       37 38.10714  1.745834 39.95238 1.704558 39.04659 38.80581 1.187327  63
QNSL      37 37.49735  1.026605 38.98942 1.406731 37.03163 38.37930 1.012102 189
XS        32 40.05631  8.171844 38.90541 2.609998 42.83728 35.13200 1.445045  74

Large Mammals
specpool(leechotu_tb, reserve_tb$reservecode2)
     Species     chao  chao.se    jack1 jack1.se    jack2     boot   boot.se   n
BME       21 29.30769 9.424882 26.53846 2.967277 30.06410 23.49887 1.6820160  13
BMNP      19 27.10000 9.200272 24.40000 3.957272 27.77778 21.44498 2.1490120  10
HSL       28 32.47187 7.150741 30.98125 1.721225 32.96246 29.39202 0.9741726 160
LL        23 23.86786 1.422226 25.89286 1.670192 24.20767 24.77311 1.2098250  28
PST       26 26.65608 1.288589 27.96825 1.391766 27.04685 27.23961 0.9590933  63
QNSL      27 27.00000 0.000000 27.00000 0.000000 23.06338 27.59689 0.7263517 189
XS        23 27.11036 4.822665 27.93243 2.205851 29.91836 25.35445 1.2855733  74
specpool(leechotu_tb, reserve_tb$reservecode)
     Species     chao  chao.se    jack1 jack1.se    jack2     boot   boot.se   n
BM        24 28.30435 4.613871 29.73913 3.052162 31.73123 26.70701 1.7293762  23
HSL       28 32.47187 7.150741 30.98125 1.721225 32.96246 29.39202 0.9741726 160
LL        23 23.86786 1.422226 25.89286 1.670192 24.20767 24.77311 1.2098250  28
PST       26 26.65608 1.288589 27.96825 1.391766 27.04685 27.23961 0.9590933  63
QNSL      27 27.00000 0.000000 27.00000 0.000000 23.06338 27.59689 0.7263517 189
XS        23 27.11036 4.822665 27.93243 2.205851 29.91836 25.35445 1.2855733  74


#### Species richness by Reserve:  iNEXT analysis
input is the Site X OTU table (e.g. leechotu5_mammals_t), which is subset into habitats by the reserve_named_mammals$reservecode or reservecode2 
```{r inext fxn reservecode, eval = FALSE}
inextcomm <- function(taxon = "mammals") { 
 # c("mammals", "lgmammals", "carnivora", "critters", "artiodactyla", "primates")

	# choose taxon subdataset for downstream analyses
	leechotu_tb <- get(str_c("leechotu5_", taxon, "_tb"))
	reserve_tb <- get(str_c("reserve_", taxon))

	# reserve subsets
	BM <- leechotu_tb %>% dplyr::filter(reserve_tb$reservecode %in% c("BM"))
	HSL <- leechotu_tb %>% dplyr::filter(reserve_tb$reservecode %in% c("HSL"))
	LL <- leechotu_tb %>% dplyr::filter(reserve_tb$reservecode %in% c("LL"))
	PST <- leechotu_tb %>% dplyr::filter(reserve_tb$reservecode %in% c("PST"))
	QNSL <- leechotu_tb %>% dplyr::filter(reserve_tb$reservecode %in% c("QNSL"))
	XS <- leechotu_tb %>% dplyr::filter(reserve_tb$reservecode %in% c("XS"))

	cname <- c("BM", "HSL", "LL", "PST", "QNSL", "XS")

	comm4inext <- matrix(c(colSums(BM), colSums(HSL), colSums(LL), colSums(PST), colSums(QNSL), colSums(XS)), ncol = 6)
	# rownames(comm4inext) <- rname
	colnames(comm4inext) <- cname
	comm4inext <- rbind(c(nrow(BM), nrow(HSL), nrow(LL), nrow(PST), nrow(QNSL), nrow(XS)), comm4inext)
	comm4inext
}

taxon <- "lgmammals"
comm4inext <- inextcomm(taxon)
   # c("mammals", "lgmammals", "carnivora", "critters", "artiodactyla", "primates")
comm4inext
confnum=0.95 # set confidence here

outcomm0 <- iNEXT(comm4inext, q=0, conf=confnum, datatype="incidence_freq")
# Hill numbers:  0 = sp richness, 1 = Shannon, 2 = inverse Simpson
sprich <- ggiNEXT(outcomm0, type=1) + ggtitle("Species richness") + labs(y="Species richness", x = "") + theme(legend.position="none")

outcomm1 <- iNEXT(comm4inext, q=1, conf=confnum, datatype="incidence_freq")
shannon <- ggiNEXT(outcomm1, type=1)  + ggtitle("Shannon diversity")  + labs(y="Shannon diversity", x = "Number of sampling units") # sample-size-based rarefaction/extrapolation curve:  Shannon

plot_grid(sprich, shannon, nrow = 2, labels = c(taxon, taxon))

outcomm2 <- iNEXT(comm4inext, q=2, conf=confnum, datatype="incidence_freq")
ggiNEXT(outcomm2, type=1) # sample-size-based rarefaction/extrapolation curve:  inverse Simpson

ggiNEXT(outcomm0, type=2) # sample completeness curve
ggiNEXT(outcomm0, type=3) # coverage-based rarefaction/extrapolation curve
# sample-size-based rarefaction/extrapolation curve:  species richness
outcomm0$DataInfo
ChaoRichness(comm4inext, datatype="incidence_freq") # same as specpool results, so i trust that we have done this correctly
ChaoShannon(comm4inext, datatype="incidence_freq")
# outcomm0$DataInfo
# outcomm0$iNextEst
# outcomm0$AsyEst
```

```{r inext fxn reservecode2, eval = FALSE}
inextcomm <- function(taxon = "mammals") { 
	# taxon == c("mammals", "lgmammals", "carnivora", "critters", "artiodactyla", "primates")

	# choose taxon subdataset for downstream analyses
	leechotu_tb <- get(str_c("leechotu5_", taxon, "_tb"))
	reserve_tb <- get(str_c("reserve_", taxon))

	# reserve subsets
	BME <- leechotu_tb %>% dplyr::filter(reserve_tb$reservecode2 %in% c("BME"))
	BMNP <- leechotu_tb %>% dplyr::filter(reserve_tb$reservecode2 %in% c("BMNP"))
	HSL <- leechotu_tb %>% dplyr::filter(reserve_tb$reservecode2 %in% c("HSL"))
	LL <- leechotu_tb %>% dplyr::filter(reserve_tb$reservecode2 %in% c("LL"))
	PST <- leechotu_tb %>% dplyr::filter(reserve_tb$reservecode2 %in% c("PST"))
	QNSL <- leechotu_tb %>% dplyr::filter(reserve_tb$reservecode2 %in% c("QNSL"))
	XS <- leechotu_tb %>% dplyr::filter(reserve_tb$reservecode2 %in% c("XS"))

	cname <- c("BME", "BMNP", "HSL", "LL", "PST", "QNSL", "XS")

	comm4inext <- matrix(c(colSums(BME), colSums(BMNP), colSums(HSL), colSums(LL), colSums(PST), colSums(QNSL), colSums(XS)), ncol = 7)
	colnames(comm4inext) <- cname
	comm4inext <- rbind(c(nrow(BME), nrow(BMNP), nrow(HSL), nrow(LL), nrow(PST), nrow(QNSL), nrow(XS)), comm4inext)
	comm4inext
}

taxon <- "critters"
comm4inext <- inextcomm(taxon)
   # c("mammals", "lgmammals", "carnivora", "critters", "artiodactyla", "primates")
comm4inext
confnum=0.95 # set confidence here

outcomm0 <- iNEXT(comm4inext, q=0, conf=confnum, datatype="incidence_freq")
# Hill numbers:  0 = sp richness, 1 = Shannon, 2 = inverse Simpson
sprich <- ggiNEXT(outcomm0, type=1) + ggtitle("Species richness") + labs(y="Species richness", x = "") + theme(legend.position="none")

outcomm1 <- iNEXT(comm4inext, q=1, conf=confnum, datatype="incidence_freq")
shannon <- ggiNEXT(outcomm1, type=1)  + ggtitle("Shannon diversity")  + labs(y="Shannon diversity", x = "Number of sampling units") # sample-size-based rarefaction/extrapolation curve:  Shannon

plot_grid(sprich, shannon, nrow = 2, labels = c(taxon, taxon))

outcomm2 <- iNEXT(comm4inext, q=2, conf=confnum, datatype="incidence_freq")
ggiNEXT(outcomm2, type=1) # sample-size-based rarefaction/extrapolation curve:  inverse Simpson

ggiNEXT(outcomm0, type=2) # sample completeness curve
ggiNEXT(outcomm0, type=3) # coverage-based rarefaction/extrapolation curve
# sample-size-based rarefaction/extrapolation curve:  species richness
outcomm0$DataInfo
ChaoRichness(comm4inext, datatype="incidence_freq") # same as specpool results, so i trust that we have done this correctly
ChaoShannon(comm4inext, datatype="incidence_freq")
# outcomm0$DataInfo
# outcomm0$iNextEst
# outcomm0$AsyEst
```

#### Community analyses

#### Run tabasco before removing zerotons and singletons (i.e. species that appear in 0 or 1 sample)
```{r}
taxon <- "mammals"
leechotu_tb <- get(str_c("leechotu5_", taxon, "_tb"))
reserve_tb <- get(str_c("reserve_", taxon))

community.jmds <- metaMDS(leechotu_tb, distance = "jaccard", trymax = 20, binary=TRUE)
community.jmds <- metaMDS(leechotu_tb, distance = "jaccard", binary = TRUE, previous.best = community.jmds)  
stressplot(community.jmds)
tabasco(leechotu_tb, use = community.jmds, labCol = reserve_tb$reservecode)
```


#### Run betapart analyses before removing zerotons and singletons
```{r}
taxon <- "mammals"
leechotu_tb <- get(str_c("leechotu5_", taxon, "_tb"))
reserve_tb <- get(str_c("reserve_", taxon))

# no nestedness 
(ALL.dist <- leechotu_tb %>% beta.multi(index.family="jac"))
# $beta.jtu
# [1] 0.9933299
# $beta.jne
# [1] 0.003278213
# $beta.jac
# [1] 0.9966081

ALL.dist <- leechotu_tb %>% beta.pair(index.family="jac")
# turnover only
ALL.dist.subset <- ALL.dist[["beta.jtu"]]
ALL.dist.jtu.jmds <- metaMDS(ALL.dist.subset)
ALL.dist.jtu.jmds <- metaMDS(ALL.dist.subset, previous.best = ALL.dist.jtu.jmds)
stressplot(ALL.dist.jtu.jmds)
# nestedness only
ALL.dist.subset <- ALL.dist[["beta.jne"]]
ALL.dist.jne.jmds <- metaMDS(ALL.dist.subset)
ALL.dist.jne.jmds <- metaMDS(ALL.dist.subset, previous.best = ALL.dist.jne.jmds)
stressplot(ALL.dist.jne.jmds)

par(mfrow=c(2,2))
plot(community.jmds, main = "All beta diversity", xlim=c(-3,3), ylim=c(-2,2))

plot(ALL.dist.jtu.jmds, main = "Turnover beta diversity only", xlim=c(-3,3), ylim=c(-2,2))

plot(ALL.dist.jne.jmds, main = "Nestedness beta diversity only", xlim=c(-3,3), ylim=c(-2,2))
par(mfrow=c(1,1))
```


#### Unconstrained ordination
#### For NMDS, mvabund, and boral analyses, remove zeroton and singleton species (these are species that appear in 0 or 1 sample). This is dfferent from the previous specnumber filtering step, where i removed samples that had 0 or 1 species only in them, which are samples that cause ordination to create strong outlier points. Such species are needed above for estimating species richness (inext, specpool) but do not provide useful information for ordination, which is driven by shared species
```{r remove zeroton and singleton species from communityB}
taxon <- "lgmammals"
leechotu_tb <- get(str_c("leechotu5_", taxon, "_tb"))
reserve_tb <- get(str_c("reserve_", taxon))

leechotu_tb <- leechotu_tb[, which(specnumber(leechotu_tb, MARGIN=2) > 1)]
```

## NMDS ordination
```{r NMDS}
### do NMDS analysis to see overall patterns ####
community.jmds <- metaMDS(leechotu_tb, distance = "jaccard", trymax = 20, binary=TRUE)
community.jmds <- metaMDS(leechotu_tb, distance = "jaccard", binary = TRUE, previous.best = community.jmds)  # doesn't converge well, with final stress > 0.20
stressplot(community.jmds)
speciesloadings <- community.jmds[["species"]] # shows species loadings
speciesloadings <- speciesloadings %>% as.data.frame() %>% rownames_to_column() 
write_tsv(speciesloadings, "analysis/speciesloadings.tsv")
```

```{r plot NMDS by reservecode}
#### plot the communities 
levels(reserve_tb$reservecode)
# "BM"   "HSL"  "LL"   "PST"  "QNSL" "XS"  

(sprichness <- specnumber(leechotu_tb, MARGIN = 1)) # number of species per site
(speciesincidence <- specnumber(leechotu_tb, MARGIN = 2)) # number of sites per species
colvec <- brewer.pal(5, "Set1")

plot(community.jmds, display="sites", type="n")

with(reserve_tb, ordiellipse(community.jmds, reservecode, cex=.5, draw="polygon", col=as.character(colvec[1]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("BM"))))

with(reserve_tb, ordiellipse(community.jmds, reservecode, cex=.5, draw="polygon", col=as.character(colvec[2]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("HSL"))))

with(reserve_tb, ordiellipse(community.jmds, reservecode, cex=.5, draw="polygon", col=as.character(colvec[3]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("LL"))))

with(reserve_tb, ordiellipse(community.jmds, reservecode, cex=.5, draw="polygon", col=as.character(colvec[3]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("PST"))))

with(reserve_tb, ordiellipse(community.jmds, reservecode, cex=.5, draw="polygon", col=as.character(colvec[4]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("QNSL"))))

with(reserve_tb, ordiellipse(community.jmds, reservecode, cex=.5, draw="polygon", col=as.character(colvec[3]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("XS"))))

# text(community.jmds, display = "sites", cex=0.7, col="blue")
text(community.jmds, display = "species", cex=0.7, col="blue")
```

```{r plot NMDS by reservecode2}
#### plot the communities 
levels(reserve_tb$reservecode2)
 # "BME"  "BMNP" "HSL"  "LL"   "PST"  "QNSL" "XS"

(sprichness <- specnumber(leechotu_tb, MARGIN = 1)) # number of species per site
(speciesincidence <- specnumber(leechotu_tb, MARGIN = 2)) # number of species per site
colvec <- brewer.pal(5, "Set1")

plot(community.jmds, display="sites", type="n")

with(reserve_tb, ordiellipse(community.jmds, reservecode2, cex=.5, draw="polygon", col=as.character(colvec[1]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("BME"))))

with(reserve_tb, ordiellipse(community.jmds, reservecode2, cex=.5, draw="polygon", col=as.character(colvec[2]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("BMNP"))))

with(reserve_tb, ordiellipse(community.jmds, reservecode2, cex=.5, draw="polygon", col=as.character(colvec[3]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("HSL"))))

with(reserve_tb, ordiellipse(community.jmds, reservecode2, cex=.5, draw="polygon", col=as.character(colvec[4]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("LL"))))

with(reserve_tb, ordiellipse(community.jmds, reservecode2, cex=.5, draw="polygon", col=as.character(colvec[1]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("PST"))))

with(reserve_tb, ordiellipse(community.jmds, reservecode2, cex=.5, draw="polygon", col=as.character(colvec[2]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("QNSL"))))

with(reserve_tb, ordiellipse(community.jmds, reservecode2, cex=.5, draw="polygon", col=as.character(colvec[3]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("XS"))))

text(community.jmds, display = "species", cex=0.7, col="blue")
```

```{r plot NMDS by reservecode3}
#### plot the communities 
levels(reserve_tb$reservecode3)
 # "BMHQNXS"  "LL"   "PST" 

(sprichness <- specnumber(leechotu_tb, MARGIN = 1)) # number of species per site
(speciesincidence <- specnumber(leechotu_tb, MARGIN = 2)) # number of species per site
colvec <- brewer.pal(5, "Set1")

plot(community.jmds, display="sites", type="n")

with(reserve_tb, ordiellipse(community.jmds, reservecode3, cex=.5, draw="polygon", col=as.character(colvec[1]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("BMHQNXS"))))

with(reserve_tb, ordiellipse(community.jmds, reservecode3, cex=.5, draw="polygon", col=as.character(colvec[4]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("LL"))))

with(reserve_tb, ordiellipse(community.jmds, reservecode3, cex=.5, draw="polygon", col=as.character(colvec[1]), alpha=20, kind="se", conf=0.95, label=TRUE, show.groups=(c("PST"))))

text(community.jmds, display = "species", cex=0.7, col="blue")
```





